\chapter*{Acknowledgments} 

I would like to take this opportunity to thank many people without whom this thesis would not have been possible.  First of all, I would like to thank my advisor Professor Veeraraghavan, who is always so brilliant and responsible. When I first came to UVa to start my graduate career, this research area was a whole new world for me. Professor Veeraraghavan was always so patient and supportive. She taught me not only knowledge in this networking area, but also the way to conduct research and tackle problems which I believe would be beneficial throughout my career and life. She helped me identify and develop a strong interest in computer networks research. Most importantly, her support and guidance helped me accumulate confidence along the way and realize what I am capable of. I would also like to thank my committee members, Professor Dugan and Professor Qi for taking the time to review my thesis and provide great suggestions.

This work was supported by NSF grants CNS-1116081,
OCI-1127340, ACI-1340910, and CNS-1405171, and the U.S.
Department of Energy (DOE) grant DE-SC0007341. I would like to thank our CC-NIE project Program Manager Kevin Thompson for giving me this opportunity to conduct networking research. 

I am fortunate to have a group of good colleagues and friends around me. I want to thank Fatemah Alali, Xiaoyu Wang, Sourav Maji, Fabrice Mizero, Gholamreza Rahimi, and Shawn Chen, they are smart, pleasant, and inspiring. We went through a lot of tough times together. Their great company has made graduate life much more fun and enjoyable. I sincerely wish them luck and success in future career, wherever they may be.

I want to thank UVA alumni students Yicheng Liang, Scott Tepsuporn and Xin Song for their contributions in this project. Without their help, these paper won't be published so successfully.

Moreover, I would like to thank our collaborators Steve Emmerson, Brian Cashman and Joseph Slezak. Your help and contributions really inspired me and played a significant role in this research project. I also would like to thank Chin Guok and AJ Ragusa for OSCARS and OESS input in Chapter.~\ref{sec:DYNES}.

Lastly, I want to thank my family without whom I cannot imagine what life would be like. Thanks for always supporting me and bringing me so many resources. I wish them good health and happiness all the time.


